%%% FORMAL DEFINITION %%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Requirements}

LDBC-SNB is designed to be flexible and to have an affordable entry point.
From small single node and in memory systems to large distributed multi-node
clusters have its own place in LDBC-SNB.  Therefore, the requirements to
fulfill for executing LDBC-SNB are limited to pure software requirements to be
able to run the tools. All the software provided by LDBC-SNB have been
developed and tested under Linux-based operating systems.

LDBC-SNB does not impose the usage of any specific type of system, as it
targets systems of different nature and characteristics, from graph databases,
graph processing frameworks and RDF systems, to traditional relational database
management systems. Consequently, any language or API capable of expressing the
proposed queries can be used. Similarly, data can be stored in the most
convenient manner the test sponsor may decide.

%, as long as it conforms with the
%execution rules. Finally, in order to have an official benchmark execution, the
%results will have to be audited and all the required information disclosed.

\section{Software and Useful links} 

\begin{itemize}
    \item \textbf{LDBC Driver 0.3 -- \url{https://github.com/ldbc/ldbc_driver}}: The driver
    responsible for executing the LDBC SNB workload.
    \item \textbf{Datagen 0.2.5 -- \url{https://github.com/ldbc/ldbc_snb_datagen}}: The data
    generator used to generate the datasets of the benchmark.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{data}
\input{workloads}
%\input{metrics}
%\input{execution}
%\input{disclosure}
%\input{auditing}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

